
## Experiment 0: Vowels

Prior to the stop analyses we scrutinezed the formant structure, F1 and F2, of 
the vowel following the coronal stops.
The purpose of this analysis was to determine if the low /a/ and /ae/ vowels 
of Spanish and English, respectively, differed from each other. 
This analysis was taken as a precautionary measure with the objective of 
determining whether or not coarticulation effects may be present in the stop 
metrics due to the possible different spectral envelopes of the next 
segment. 

The F1 and F2 data were analayzed using separate Bayesian multilevel models.
The criterion (F1 or F2) were standardized and modeled as a function of 
language (English, Spanish), phoneme (/d/, /t/), and item repetition. 
Language and phoneme were sum coded (-1, 1). 
The random effects structure included by-subject intercepts with 
random slopes for phoneme and item repetition, as well as by-item intercepts 
with a random slope for item repetition. 
The model included weakly informative priors with the mean centered at 0 and 
a standard deviation of 2. 

The model suggested weak evidence for a language effect on F1 
`r print_model_param(all_models$mod_f1_mono_full, "language_sum")`
and F2 
`r print_model_param(all_models$mod_f2_mono_full, "language_sum")`. 
Fifteen and eighteen percent of the HDI's fell within the specified ROPE for 
F1 and F2, respectively. 
The probability that &beta; was greater than 0 was .979 for F1 and .958 for F2. 
Together, the point estimates suggest the spectral centroid of the Spanish 
vowel was slightly higher and more posterior with respect to that of the 
English vowel.
There was no evidence for a phoneme effect on F1 
`r print_model_param(all_models$mod_f1_mono_full, "phon_sum")`, 
nor on F2 
`r print_model_param(all_models$mod_f2_mono_full, "phon_sum")`. 
In both cases 100% of the HDI fell within the ROPE and the probability that 
&beta; was less than 0 was .717 for F1 and .555 for F2. 
Given the possibility of a small effect of language on the spectal envelope, 
standardized F1 and F2 were used in subsequent analyses of the coronal stops 
to control for any coarticulatory effects on the burst. 
Figure\ \@ref(fig:plot-monolingual-vowels) plots the F1 &#215; F2 data and 
Figure\ \@ref(fig:plot-monolingual-vowels-summary) plots the model summary. 
A complete summary of the F1 and F2 models is available in Table 1. 


(ref:plot-monolingual-vowels) F1 &#215; F2 of /a/ from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 95% and 80% credible intervals.

```{r, 'plot-monolingual-vowels', echo=FALSE, fig.cap="(ref:plot-monolingual-vowels)"}
knitr::include_graphics(
  here("figs", "vowel_all_metrics.pdf")
)
```


(ref:plot-monolingual-vowels-summary) Posterior medians &plusmn; 95% and 66% credible intervals for F1 and F2 from monolingual speaker data.

```{r, 'plot-monolingual-vowels-summary', echo=FALSE, fig.cap="(ref:plot-monolingual-vowels-summary)"}
knitr::include_graphics(
  here("figs", "vowel_summary.pdf")
)
```




```{r, 'vowel-table', echo=F, fig.cap="(ref:vowel-table)"}
my_border <- officer::fp_border(width = 1.5)

read_csv(here("data", "tidy", "table_vowel_model_summary.csv")) %>% 
  flextable() %>%
  width(., j = c(2, 3, 4, 5, 6), width = c(1.2, 0.7, 1.25, 0.7, 0.7)) %>% 
  align(., i = NULL, j = 4, align = "right", part = "all") %>% 
  padding(., i = NULL,  padding.bottom = 0, padding.top = 0, part = "body") %>%
  padding(., i = 4,  padding.bottom = 4, padding.top = 0, part = "body") %>%
  font(., fontname = "Times", part = "all") %>%
  fontsize(., size = 11) %>%
  border_remove(.) %>%
  border(., part = "header",
            border.top = my_border,
            border.bottom = my_border) %>%
  hline_bottom(., part = "body", border = my_border) %>%
  set_caption(., caption = "*Table A1*: Model summary for F1 and F2 as a function of language (English, Spanish), phoneme (/d/, /t/), and item repetition for monolingual data. The percentage of the HDI contained within the ROPE is based on an effect size of &plusmn; 0.1.")
```
