## Appendix C: Supplementary analyses (bilinguals)

In this section we present additional information from the analyses of the bilingual data. 
@tbl-bi includes the full output of the model summary from the primary multivariate analysis (Note: this table is equivalent to the forest plot in @fig-bilinguals). 

```{r}
#| label: tbl-bi
#| results: asis
bi_table %>%  
  mutate(
    Parameter = str_replace(Parameter, "x", "&#215;"), 
    Metric = case_when(
    identifier == "VOT_Intercept" ~ "VOT", 
    identifier == "COG_Intercept" ~ "COG", 
    identifier == "KT_Intercept" ~ "KT", 
    identifier == "RI_Intercept" ~ "RI", 
    identifier == "SD_Intercept" ~ "SD", 
    identifier == "SK_Intercept" ~ "SK", 
    TRUE ~ " ")) %>% 
  select(-identifier) %>% 
  knitr::kable(format = "markdown", 
    align = c("l", "l", "r", "r", "r", "r"), 
    caption = "Model summary for VOT and burst metrics as a function of language (English, Spanish), phoneme (/d/, /t/), F1, F2, and item repetition for bilingual coronal stops. The percentage of the HDI contained within the ROPE is based on an effect size of &plusmn;0.1. {#tbl-bi}")
```

<br>

@fig-bilinguals-d-t shows the comparison of English /d/ and Spanish /t/ for voice-onset time. 
One can observe that there is not a meaningful difference between the short-lag segments. 


```{r}
#| label: fig-bilinguals-d-t
#| fig-cap: "Posterior distribution comparing the short-lag stops, English /d/ and Spanish /t/. The white point represents the posterior mean &pm;95% HDI and the grey region represents the ROPE (&pm;0.1)."
knitr::include_graphics(
  here("figs", "bi_post_hoc_dt.pdf")
)
```


@fig-bi-post-hoc-sm and @tbl-bi-post-hoc-sm present the post-hoc analyses of within-language comparisons, that is, /d/ vs. /t/. 

```{r}
#| label: fig-bi-post-hoc-sm
#| fig-cap: "Posterior distributions of difference estimates between voiced and voiceless stops for Spanish and English. The point represents the posterior mean difference &pm;95% HDI of the posterior of the voiceless segment subtracted from the posterior of the voiced. The light blue region represents the ROPE (&pm;0.1)."
knitr::include_graphics(
  here("figs", "bi_post_hoc_sm.pdf")
)
```


```{r}
#| label: tbl-bi-post-hoc-sm
#| results: asis
post_hoc %>% 
  filter(Metric != "VOT", Parameter == "bi_dt_win_lang") %>% 
  select(-identifier, -Parameter) %>% 
  separate(Metric, into = c("Language", "Metric"), sep = "_") %>% 
  arrange(desc(Language)) %>% 
  mutate(Language = str_to_title(Language), 
    Language = case_when(
      Metric != "COG" & Language == "Spanish" ~ " ", 
      Metric != "COG" & Language == "English" ~ " ", 
      TRUE ~ Language
    )) %>% 
    knitr::kable(format = "markdown", 
    align = c("l", "l", "r", "r", "r", "r"), 
    caption = "Post-hoc comparisons of within-language voiced-voiceless pairs for each burst metric. Each point represents the posterior mean difference &pm;95% HDI of the posterior of the voiceless segment subtracted from the posterior of the voiced segment. {#tbl-bi-post-hoc-sm}")
```
