
## Appendix A: Vowels

Prior to the stop analyses we scrutinized the formant structure, F1 and F2, of  the vowel following the coronal stops.
The purpose of this analysis was to determine if the low /a/ and /&aelig;/ vowels of Spanish and English, respectively, differed from each other.
This analysis was taken as a precautionary measure with the objective of determining whether or not coarticulation effects may be present in the stop metrics due to the possible different spectral envelopes of the next segment.

The F1 and F2 data were analyzed using separate Bayesian multilevel models.
The criterion (F1 or F2) were standardized and modeled as a function of language (English, Spanish), phoneme (/d/, /t/), and item repetition.
Language and phoneme were sum coded (-1, 1). 
The random effects structure included by-subject intercepts with random slopes for phoneme and item repetition, as well as by-item intercepts with a random slope for item repetition. 
The model included weakly informative priors with the mean centered at 0 and a standard deviation of 2.

The model suggested weak evidence for a language effect on F1 `r report_posterior(f1_f2_table, param = "Language", metric = "F1")` and F2
`r report_posterior(f1_f2_table, param = "Language", metric = "F2")`. 
Together, the point estimates suggest the spectral centroid of the Spanish vowel was slightly higher and more posterior with respect to that of the English vowel.
There was no evidence for a phoneme effect on F1 `r report_posterior(f1_f2_table, param = "Voicing", metric = "F1")`, nor on F2 `r report_posterior(f1_f2_table, param = "Voicing", metric = "F2")`.
Given the possibility of a small effect of language on the spectral envelope,  standardized F1 and F2 were used in subsequent analyses of the coronal stops to control for any coarticulatory effects on the burst.
@fig-monolingual-vowels plots the F1 &#215; F2 data and @fig-monolingual-vowels-summary plots the model summary. 
A complete summary of the F1 and F2 models is available in @tbl-vowels.

```{r}
#| label: fig-monolingual-vowels
#| fig-cap: "F1 &#215; F2 of /a/ from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn;95% and 80% credible intervals."
include_graphics(here("figs", "vowel_all_metrics_marginal.pdf"))
```

```{r}
#| label: fig-monolingual-vowels-summary
#| fig-cap: "Posterior medians &plusmn;95% and 66% credible intervals for F1 and F2 from monolingual speaker data."
include_graphics(here("figs", "vowel_summary.pdf"))
```



```{r}
#| label: tbl-vowels
#| tbl-cap: "Model summary for F1 and F2 as a function of language (English, Spanish), phoneme (/d/, /t/), and item repetition for monolingual data. The percentage of the HDI contained within the ROPE is based on an effect size of &plusmn;0.1. {#tbl-vowels}"
f1_f2_table %>% 
  mutate(Metric = case_when(
      identifier == "F1_Intercept" ~ "F1", 
      identifier == "F2_Intercept" ~ "F2", 
      TRUE ~ " ")) %>% 
  select(-identifier) %>% 
  kable(format = "markdown", 
    align = c("l", "l", "r", "r", "r", "r"))
```
