---
title             : "The title"
shorttitle        : "Title"

author: 
  - name          : "Miquel Simonet"
    affiliation   : "1"
    corresponding : yes
    address       : "1423 E. University Blvd. Rm. 594 Modern Languages Building, Tucson, Arizona 85721"
    email         : "simonet@email.arizona.edu"
  - name          : "Joseph V. Casillas"
    affilitation  : "2"
  - name          : "Alex Aldrich"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Arizona"
  - id            : "2"
    institution   : "Rutgers University"
  - id            : "3"
    institution   : "Concordia College"

authornote: |
  Add complete departmental affiliations for each author here. Each new line 
  herein must be indented, like this line.

  Enter author note here.

abstract: |
  Here. 


keywords          : "Spanish, Coronal stops, Spectral moments"
wordcount         : ""

bibliography      : 
  - "r-references.bib"
  - "stats-references.bib"

appendix:
  - "appendix-a.Rmd"
  - "appendix-b.Rmd"
  - "appendix-c.Rmd"
  - "appendix-d.Rmd"
  - "appendix-e.Rmd"
  - "appendix-f.Rmd"

floatsintext      : yes
figurelist        : yes
tablelist         : yes
footnotelist      : yes
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r 'doc-libs', include = FALSE}
library("papaja")
library("officer")
library("brms")
library("knitr")
```

```{r 'setup', include=FALSE}
set.seed(42)
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      cache.extra = knitr::rand_seed)
options(width=80)
```

```{r, 'load-all'}
source(here::here("scripts", "r", "production", "99_small_data.R"))
```

# Method

## Participants

The data include `r exp_n` participants from 3 populations: monolingual 
English speakers, monolingual Spanish speakers, and bilingual Spanish-English 
speakers.
All participants were females between the ages of 18 and 23. 
The monolingual English speakers were recorded in English and the monolingual 
Spanish speakers were recorded in Spanish. 
The Spanish-English bilinguals were recorded in both of their languages.

**Monolingual English speakers.** The study includes `r mono_en_n` monolingual 
English speakers.
They were undergraduate students at the University of Arizona, born and raised 
in the US Southwest. 
The English speakers were functionally monolingual, though they reported having 
taken introductory Spanish courses. 
They were not able to maintain a basic conversation in Spanish.
All of the participants in this group reported English as their native 
language and verified not having been exposed to any other languages while 
growing up.


**Monolingual Spanish speakers.** The monolingual Spanish group comprised 
`r mono_sp_n` speakers that were recruited from the 
*Universitat de les  Illes Balears* campus community and were born and raised 
on the island of Majorca, Spain.
They reported that, although they had studied some English in Spain, they were 
not able to maintain a basic conversation in this language. 
The participants of this group also speak Catalan. 
Importantly, there are no reported differences in the phonetic realization of 
voice timing between the Spanish and Catalan, nor are there place difference 
between the coronal stops. 

**Bilingual speakers.** The English-Spanish bilinguals (n = 
`r bi_coronals_n + bi_poa_n`) came from Southern Arizona and Northern Mexico.
There are two samples from this population. 
The coronal dataset includes `r bi_coronals_n` speakers and the bilabial 
dataset includes `r bi_poa_n` speakers. 
The Spanish-English bilinguals were undergraduate students at the University 
of Arizona in Tucson, Arizona. 
The bilinguals were brought up by Spanish-speaking families and were schooled 
mostly in English. 
They reported using English and Spanish daily, both in the classroom as well 
as with their friends and relatives.


## Metrics

F1/F2, voice onset time, relative intensity, center of gravity, standard 
deviation, skewness, kurtosis

## Procedure

Decide if we will present 3 separate experiments with 3 different methods 
sections.

## Statistical analyses

All analyses were conducted in `R` [@R-base, `r r_version`]. 
We use Bayesian multilevel models fitted in `Stan` using `brms` 
[@R-brms_a; @R-brms_b, `r brms_version`].
Bayesian Data Analysis (BDA) is an alternative to frequentist statistical 
analysis.^[See @van2014bayesian and @vasishth2018bayesian for tutorials and in 
depth explanations related to BDA in the psychological and speech sciences.] 
For all models, the criterion was standardized, or converted to z-scores, 
in order to facilitate comparability between metrics. 
Continuous predictors were also standardized and categorical predictors were 
sum-to-zero coded. 
Thus for all models the intercept represents the outcome variable at the 
grand mean in standardized units. 
We used regularizing, weakly informative priors in all models (specifics below) 
with 4,000 iterations (2,000 warm-up) running on 16 processing cores. 
We quantify our uncertainty regarding a given effect by reporting point 
estimates derived from the posterior predictive distribution, including the 
95% Highest Density Credible Intervals (HDI). 
Additionally, we assume a negligible effect size of &plusmn; 0.1 
[@cohen1988statistical; @cohen2013statistical; @kruschke2018rejecting] 
in order to establish a Region of Practical Equivalence (ROPE), for which we 
assess the proportion of the HDI that falls within this interval.
Finally, we report the Maximum Probability of Effect (MPE), or the Probability 
of Direction, as the proportion of the posterior distribution that is of the 
median's sign. 
We assume there to be compelling evidence for a given effect when the HDI of 
the posterior distribution does not contain 0 nor fall within the ROPE by a 
reasonable margin and the MPE is close to 1. 




# Results

The results are divided into 3 subsections dealing with (1) monolingual data, 
(2) bilingual data, and (3) POA data. 
Each subsection presents the results of 6 metrics: VOT, relative intensity, 
center of gravity, kurtosis, standard deviation, and skewness. 
We report only relevant effects. 
Please see the supplementary materials (Appendices A-D) for complete model 
summaries. 

## Experiment 1: Monolinguals

We modeled VOT and the burst metrics as a function of language (English, 
Spanish), phoneme (/d/, /t/), standardized F1 and F2, and item repetitions. 
The model used regularizing, weakly informative priors \citep{Gelman_2017, 
vasishth2018bayesian}.
Specifically, all parameters were assumed to be distributed as normal with a 
standard deviation of 5, i.e., $Normal(\mu = 0$, $\sigma = 5)$. 
The standard deviation parameters for random effects and the model residual 
error (sigma) were truncated to exclude negative values. 
Figure\ \@ref(fig:plot-monolinguals) plots VOT and the burst metrics as a 
function of language (English, Spanish) and phoneme (/d/, /t/).
For all plots the y-axis represents the outcome variable in standardized units. 
The x-axis indicates the language and voiced/voiceless pairs are represented 
by color. 

(ref:plot-monolinguals) VOT and burst metrics of coronal stops (/d/, /t/) from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-monolinguals', echo=FALSE, fig.cap="(ref:plot-monolinguals)"}
knitr::include_graphics(
  here("figs", "mono_all_metrics.pdf")
)
```

**Voice-onset time**. English stops had higher VOT than Spanish stops 
`r report_posterior(mono_table, param = "Group", metric = "VOT")` and the 
voiceless segments had higher VOT than the voiced segments 
`r report_posterior(mono_table, param = "Phoneme", metric = "VOT")`. 
The HDI's for language and phoneme were completely outside the ROPE and the 
MPE was 100%.
Thus there was compelling evidence for an effect of both language and phoneme, 
as well as a moderate interaction between the two predictors
`r report_posterior(mono_table, param = "Group x Phoneme", metric = "VOT")`, 
such that the voicing difference between Spanish coronals was slightly larger 
than that of the English coronals.
Additionally, we compared the short-lag stops of each language, English /d/ 
and Spanish /t/, and found no evidence that the segments differed from each 
other REPORT HERE MAYBE.
The VOT data is plotted in the first panel of Figure 
\@ref(fig:plot-monolinguals). 
The complete model summary and the short-lag stop comparison are available in 
Table \@ref(tab:mono-table) and Figure \@ref(fig:plot-monolinguals-d-t) of the 
supplementary materials.





**Relative intensity**. The relative intensity data is plotted in the top 
middle panel of Figure \@ref(fig:plot-monolinguals). 
English and Spanish stops differed very little with regard to relative 
intensity. 
The model provided no compelling evidence for a group effect
`r report_posterior(mono_table, param = "Group", metric = "RI")`, nor for 
phoneme 
`r report_posterior(mono_table, param = "Phoneme", metric = "RI")` or vowel 
height effects 
`r report_posterior(mono_table, param = "F1", metric = "RI")`. 
Frontness of the following vowel did modulate relative intensity of the burst
`r report_posterior(mono_table, param = "F2", metric = "RI")` such that 
higher F2 values were associated with lower relative intensity. 
Finally, there was no compelling evidence for a language &#215; phoneme 
interaction
`r report_posterior(mono_table, param = "Group x Phoneme", metric = "RI")`. 






**Center of gravity.** The center of gravity (COG) data is plotted in the 
third panel of Figure \@ref(fig:plot-monolinguals). 
English stops had a higher COG than Spanish stops
`r report_posterior(mono_table, param = "Group", metric = "COG")`, and the 
voiceless segments had a higher COG than the voiced segments 
`r report_posterior(mono_table, param = "Phoneme", metric = "COG")`. 
In both cases the HDI's were completely outside the ROPE and the MPE was 100%.
This suggests compelling evidence for an effect of both language and phoneme. 
The two predictors did not interact 
`r report_posterior(mono_table, param = "Group x Phoneme", metric = "COG")`, 
nor did F1 
`r report_posterior(mono_table, param = "F1", metric = "COG")` or F2 
`r report_posterior(mono_table, param = "F2", metric = "COG")` have any 
influence on COG. 
 






**Kurtosis.** The kurtosis data is plotted in the second row, first column 
of Figure \@ref(fig:plot-monolinguals).
English stop bursts had a lower kurtosis with regard to Spanish stop bursts 
`r report_posterior(mono_table, param = "Group", metric = "KT")`, and the 
voiced segments presumably had a higher kurtosis than the voiceled segments 
`r report_posterior(mono_table, param = "Phoneme", metric = "KT")`, though 
there was evidence of a language &times; phoneme interaction 
`r report_posterior(mono_table, param = "Group x Phoneme", metric = "KT")`. 
<!--
need to add post hoc comparisons here
-->
Specifically, kurtosis was higher in the voiced stop bursts of Spanish (XXX), 
but there was no evidence of a voicing difference in the English data (XXX). 
Neither F1
`r report_posterior(mono_table, param = "F1", metric = "KT")` nor F2 
`r report_posterior(mono_table, param = "F2", metric = "KT")` had any 
influence on kurtosis.   






**Standard deviation.**

`r report_posterior(mono_table, param = "Group", metric = "SD")`  

`r report_posterior(mono_table, param = "Phoneme", metric = "SD")`  

`r report_posterior(mono_table, param = "F1", metric = "SD")`  

`r report_posterior(mono_table, param = "F2", metric = "SD")`  

`r report_posterior(mono_table, param = "Group x Phoneme", metric = "SD")`





**Skewness.**

`r report_posterior(mono_table, param = "Group", metric = "SK")`  

`r report_posterior(mono_table, param = "Phoneme", metric = "SK")`  

`r report_posterior(mono_table, param = "F1", metric = "SK")`  

`r report_posterior(mono_table, param = "F2", metric = "SK")`  

`r report_posterior(mono_table, param = "Group x Phoneme", metric = "SK")`


(ref:plot-monolinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of monolingual coronal stops.

```{r, 'plot-monolinguals-summary', echo=FALSE, fig.cap="(ref:plot-monolinguals-summary)"}
knitr::include_graphics(
  here("figs", "mono_summary.pdf")
)
```




### Interim discussion

General summary and patterns. 
ANything noteworthy (i.e. VOT of short-lag stops, F1 in RI, 
KT interaction, etc.)







## Experiment 2: Bilinguals

Model info here. 

(ref:plot-bilinguals) VOT and burst metrics of coronal stops (/d/, /t/) from bilingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-bilinguals', echo=FALSE, fig.cap="(ref:plot-bilinguals)"}
knitr::include_graphics(
  here("figs", "bi_all_metrics.pdf")
)
```

**Voice-onset time.**

`r report_posterior(bi_table, param = "Language", metric = "VOT")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "VOT")`  

`r report_posterior(bi_table, param = "F1", metric = "VOT")`  

`r report_posterior(bi_table, param = "F2", metric = "VOT")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "VOT")`

**Relative intensity.**

`r report_posterior(bi_table, param = "Language", metric = "RI")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "RI")`  

`r report_posterior(bi_table, param = "F1", metric = "RI")`  

`r report_posterior(bi_table, param = "F2", metric = "RI")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "RI")`

**Center of gravity.**

`r report_posterior(bi_table, param = "Language", metric = "COG")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "COG")`  

`r report_posterior(bi_table, param = "F1", metric = "COG")`  

`r report_posterior(bi_table, param = "F2", metric = "COG")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "COG")`

**Kurtosis.**

`r report_posterior(bi_table, param = "Language", metric = "KT")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "KT")`  

`r report_posterior(bi_table, param = "F1", metric = "KT")`  

`r report_posterior(bi_table, param = "F2", metric = "KT")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "KT")`


**Standard deviation.**

`r report_posterior(bi_table, param = "Language", metric = "SD")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "SD")`  

`r report_posterior(bi_table, param = "F1", metric = "SD")`  

`r report_posterior(bi_table, param = "F2", metric = "SD")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "SD")`

**Skewness.**

`r report_posterior(bi_table, param = "Language", metric = "SK")`  

`r report_posterior(bi_table, param = "Phoneme", metric = "SK")`  

`r report_posterior(bi_table, param = "F1", metric = "SK")`  

`r report_posterior(bi_table, param = "F2", metric = "SK")`  

`r report_posterior(bi_table, param = "Language x Phoneme", metric = "SK")`



(ref:plot-bilinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of bilingual coronal stops.

```{r, 'plot-bilinguals-summary', echo=FALSE, fig.cap="(ref:plot-bilinguals-summary)"}
knitr::include_graphics(
  here("figs", "bi_summary.pdf")
)
```













## Experiment 3: Bilingual POA data

Model info here. 


(ref:plot-poa-bilinguals) VOT and burst metrics of voiceless stops from bilingual speakers as a function of language (English, Spanish), place of articulation (Coronal, Bilabial). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-poa-bilinguals', echo=FALSE, fig.cap="(ref:plot-poa-bilinguals)"}
knitr::include_graphics(
  here("figs", "poa_all_metrics.pdf")
)
```

**Voice-onset time.**

`r report_posterior(bi_poa_table, param = "Language", metric = "VOT")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "VOT")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "VOT")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "VOT")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "VOT")`

**Relative intensity.**

`r report_posterior(bi_poa_table, param = "Language", metric = "RI")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "RI")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "RI")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "RI")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "RI")`

**Center of gravity.**

`r report_posterior(bi_poa_table, param = "Language", metric = "COG")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "COG")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "COG")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "COG")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "COG")`

**Kurtosis.**

`r report_posterior(bi_poa_table, param = "Language", metric = "KT")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "KT")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "KT")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "KT")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "KT")`

**Standard deviation.**

`r report_posterior(bi_poa_table, param = "Language", metric = "SD")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "SD")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "SD")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "SD")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "SD")`

**Skewness.**

`r report_posterior(bi_poa_table, param = "Language", metric = "SK")`  

`r report_posterior(bi_poa_table, param = "Place", metric = "SK")`  

`r report_posterior(bi_poa_table, param = "F1", metric = "SK")`  

`r report_posterior(bi_poa_table, param = "F2", metric = "SK")`  

`r report_posterior(bi_poa_table, param = "Language x Place", metric = "SK")`




(ref:plot-poa-bilinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of bilingual voiceless bilabial and coronal stops.

```{r, 'plot-poa-bilinguals-summary', echo=FALSE, fig.cap="plot-poa-bilinguals-summary"}
knitr::include_graphics(
  here("figs", "poa_summary.pdf")
)
```








# Discussion

## Summary of findings

## Interpretation and implications


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
