---
author: 
  - name          : "Miquel Simonet"
    affiliation   : "1"
    corresponding : yes
    address       : "1423 E. University Blvd. Rm. 594 Modern Languages Building, Tucson, Arizona 85721"
    email         : "simonet@email.arizona.edu"
  - name          : "Joseph V. Casillas"
    affilitation  : "2"
  - name          : "Alex Aldrich"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Arizona"
  - id            : "2"
    institution   : "Rutgers University"
  - id            : "3"
    institution   : "Concordia College"

authornote: |
  Add complete departmental affiliations for each author here. Each new line 
  herein must be indented, like this line.

  Enter author note here.

abstract: |
  Here. 


keywords          : "Spanish, Coronal stops, Spectral moments"
wordcount         : ""

bibliography      : 
  - "r-references.bib"
  - "stats-references.bib"

appendix:
  - "appendix-a.Rmd"
  - "appendix-b.Rmd"
  - "appendix-c.Rmd"
  - "appendix-d.Rmd"
  - "appendix-e.Rmd"
  - "appendix-f.Rmd"

floatsintext      : yes
figurelist        : yes
tablelist         : yes
footnotelist      : yes
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r 'doc-libs', include = FALSE}
library("papaja")
library("officer")
library("flextable")
library("brms")
```

```{r 'setup', include=FALSE}
set.seed(42)
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      cache.extra = knitr::rand_seed)
options(width=80)
```

```{r, 'load-all'}
source(here::here("scripts", "r", "production", "99_small_data.R"))
```



# Method

## Participants

The data include `r exp_n` participants from 3 populations: monolingual 
English speakers, monolingual Spanish speakers, and bilingual Spanish-English 
speakers.
All participants were females between the ages of 18 and 23. 
The monolingual English speakers were recorded in English and the monolingual 
Spanish speakers were recorded in Spanish. 
The Spanish-English bilinguals were recorded in both of their languages.

**Monolingual English speakers.** The study includes `r mono_en_n` monolingual 
English speakers.
They were undergraduate students at the University of Arizona, born and raised 
in the US Southwest. 
The English speakers were functionally monolingual, though they reported having 
taken introductory Spanish courses. 
They were not able to maintain a basic conversation in Spanish.
All of the participants in this group reported English as their native 
language and verified not having been exposed to any other languages while 
growing up.


**Monolingual Spanish speakers.** The monolingual Spanish group comprised 
`r mono_sp_n` speakers that were recruited from the 
*Universitat de les  Illes Balears* campus community and were born and raised 
on the island of Majorca, Spain.
They reported that, although they had studied some English in Spain, they were 
not able to maintain a basic conversation in this language. 
The participants of this group also speak Catalan. 
Importantly, there are no reported differences in the phonetic realization of 
voice timing between the Spanish and Catalan, nor are there place difference 
between the coronal stops. 

**Bilingual speakers.** The English-Spanish bilinguals (n = 
`r bi_coronals_n + bi_poa_n`) came from Southern Arizona and Northern Mexico.
There are two samples from this population. 
The coronal dataset includes `r bi_coronals_n` speakers and the bilabial 
dataset includes `r bi_poa_n` speakers. 
The Spanish-English bilinguals were undergraduate students at the University 
of Arizona in Tucson, Arizona. 
The bilinguals were brought up by Spanish-speaking families and were schooled 
mostly in English. 
They reported using English and Spanish daily, both in the classroom as well 
as with their friends and relatives.


## Metrics

F1/F2, voice onset time, relative intensity, center of gravity, standard 
deviation, skewness, kurtosis

## Procedure

Decide if we will present 3 separate experiments with 3 different methods 
sections.

## Statistical analyses

All analyses were conducted in `R` [@R-base, `r r_version`]. 
We use Bayesian multilevel models fitted in `Stan` using `brms` 
[@R-brms_a; @R-brms_b, `r brms_version`].
Bayesian Data Analysis (BDA) is an alternative to frequentist statistical 
analysis.^[See @van2014bayesian and @vasishth2018bayesian for tutorials and in 
depth explanations related to BDA in the psychological and speech sciences.] 
For all models, the criterion was standardized, or converted to z-scores, 
in order to facilitate comparability between metrics. 
Continuous predictors were also standardized and categorical predictors were 
sum-to-zero coded. 
Thus for all models the intercept represents the outcome variable at the 
grand mean in standardized units. 
We used regularizing, weakly informative priors in all models (specifics below) 
with 4,000 iterations (2,000 warm-up) running on 16 processing cores. 
We quantify our uncertainty regarding a given effect by reporting point 
estimates derived from the posterior predictive distribution, including the 
95% Highest Density Credible Intervals (HDI). 
Additionally, we assume a negligible effect size of &plusmn; 0.1 
[@cohen1988statistical; @cohen2013statistical; @kruschke2018rejecting] 
in order to establish a Region of Practical Equivalence (ROPE), for which we 
assess the proportion of the HDI that falls within this interval.
Finally, we report the Maximum Probability of Effect (MPE), or the Probability 
of Direction, as the proportion of the posterior distribution that is of the 
median's sign. 
We assume there to be compelling evidence for a given effect when the HDI of 
the posterior distribution does not contain 0 nor fall within the ROPE by a 
reasonable margin and the MPE is close to 1. 




# Results

The results are divided into 3 subsections dealing with (1) monolingual data, 
(2) bilingual data, and (3) POA data. 
Each subsection presents the results of 6 metrics: VOT, relative intensity, 
center of gravity, kurtosis, standard deviation, and skewness. 
We report only relevant effects. 
Please see the supplementary materials (Appendices A-D) for complete model 
summaries. 

## Experiment 1: Monolinguals

We modeled VOT and the burst metrics as a function of language (English, 
Spanish), phoneme (/d/, /t/), standardized F1 and F2, and item repetitions. 
The model used regularizing, weakly informative priors \citep{Gelman_2017, 
vasishth2018bayesian}.
Specifically, all parameters were assumed to be distributed as normal with a 
standard deviation of 5, i.e., $Normal(\mu = 0$, $\sigma = 5)$. 
The standard deviation parameters for random effects and the model residual 
error (sigma) were truncated to exclude negative values. 
Figure\ \@ref(fig:plot-monolinguals) plots VOT and the burst metrics as a 
function of language (English, Spanish) and phoneme (/d/, /t/).
For all plots the y-axis represents the outcome variable in standardized units. 
The x-axis indicates the language and voiced/voiceless pairs are represented 
by color. 

(ref:plot-monolinguals) VOT and burst metrics of coronal stops (/d/, /t/) from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-monolinguals', echo=FALSE, fig.cap="(ref:plot-monolinguals)"}
knitr::include_graphics(
  here("figs", "mono_all_metrics.pdf")
)
```

**Voice-onset time**. English stops had higher VOT than Spanish stops 
`r print_model_param(all_models$mod_coronals_vot_mono, "group_sum")` and the 
voiceless segments had higher VOT than the voiceless segments 
`r print_model_param(all_models$mod_coronals_vot_mono, "phon_sum")`. 
For both language and phoneme none of the HDI fell within the ROPE and the MPE 
was 100%. 
Thus there was compelling evidence for an effect of language and phoneme, as 
well as a moderate interaction between the two predictors 
`r print_model_param(all_models$mod_coronals_vot_mono, "group_sum:phon_sum")`, 
such that the voicing difference between Spanish coronals was larger than that 
of the English coronals.
For the language &#215; phoneme interaction .364 of the HDI was within the 
ROPE and the MPE was .998. 
The VOT data is plotted in first panel of Figure\ \@ref(fig:plot-monolinguals) 
and the complete model summary is available in Appendix B.
<!--
\ \@ref(mono-table).
-->

**Relative intensity**. 

`r print_model_param(all_models$mod_coronals_ri_mono, "group_sum")`  

`r print_model_param(all_models$mod_coronals_ri_mono, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_ri_mono, "f1_std")`  

`r print_model_param(all_models$mod_coronals_ri_mono, "f2_std")`  

`r print_model_param(all_models$mod_coronals_ri_mono, "group_sum:phon_sum")`

**Center of gravity.**

`r print_model_param(all_models$mod_coronals_cog_mono, "group_sum")`  

`r print_model_param(all_models$mod_coronals_cog_mono, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_cog_mono, "f1_std")`  

`r print_model_param(all_models$mod_coronals_cog_mono, "f2_std")`  

`r print_model_param(all_models$mod_coronals_cog_mono, "group_sum:phon_sum")`

**Kurtosis.**

`r print_model_param(all_models$mod_coronals_kt_mono, "group_sum")`  

`r print_model_param(all_models$mod_coronals_kt_mono, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_kt_mono, "f1_std")`  

`r print_model_param(all_models$mod_coronals_kt_mono, "f2_std")`  

`r print_model_param(all_models$mod_coronals_kt_mono, "group_sum:phon_sum")`

**Standard deviation.**

`r print_model_param(all_models$mod_coronals_sd_mono, "group_sum")`  

`r print_model_param(all_models$mod_coronals_sd_mono, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_sd_mono, "f1_std")`  

`r print_model_param(all_models$mod_coronals_sd_mono, "f2_std")`  

`r print_model_param(all_models$mod_coronals_sd_mono, "group_sum:phon_sum")`

**Skewness.**

`r print_model_param(all_models$mod_coronals_sk_mono, "group_sum")`  

`r print_model_param(all_models$mod_coronals_sk_mono, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_sk_mono, "f1_std")`  

`r print_model_param(all_models$mod_coronals_sk_mono, "f2_std")`  

`r print_model_param(all_models$mod_coronals_sk_mono, "group_sum:phon_sum")`


(ref:plot-monolinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of monolingual coronal stops.

```{r, 'plot-monolinguals-summary', echo=FALSE, fig.cap="(ref:plot-monolinguals-summary)"}
knitr::include_graphics(
  here("figs", "mono_summary.pdf")
)
```













## Experiment 2: Bilinguals

Model info here. 

(ref:plot-bilinguals) VOT and burst metrics of coronal stops (/d/, /t/) from bilingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-bilinguals', echo=FALSE, fig.cap="(ref:plot-bilinguals)"}
knitr::include_graphics(
  here("figs", "bi_all_metrics.pdf")
)
```

**Voice-onset time.**

`r print_model_param(all_models$mod_coronals_vot_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_vot_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_vot_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_vot_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_vot_bi, "language_sum:phon_sum")`

**Relative intensity.**

`r print_model_param(all_models$mod_coronals_ri_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_ri_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_ri_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_ri_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_ri_bi, "language_sum:phon_sum")`

**Center of gravity.**

`r print_model_param(all_models$mod_coronals_cog_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_cog_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_cog_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_cog_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_cog_bi, "language_sum:phon_sum")`

**Kurtosis.**

`r print_model_param(all_models$mod_coronals_kt_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_kt_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_kt_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_kt_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_kt_bi, "language_sum:phon_sum")`

**Standard deviation.**

`r print_model_param(all_models$mod_coronals_sd_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_sd_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_sd_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_sd_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_sd_bi, "language_sum:phon_sum")`

**Skewness.**

`r print_model_param(all_models$mod_coronals_sk_bi, "language_sum")`  

`r print_model_param(all_models$mod_coronals_sk_bi, "phon_sum")`  

`r print_model_param(all_models$mod_coronals_sk_bi, "f1_std")`  

`r print_model_param(all_models$mod_coronals_sk_bi, "f2_std")`  

`r print_model_param(all_models$mod_coronals_sk_bi, "language_sum:phon_sum")`



(ref:plot-bilinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of bilingual coronal stops.

```{r, 'plot-bilinguals-summary', echo=FALSE, fig.cap="(ref:plot-bilinguals-summary)"}
knitr::include_graphics(
  here("figs", "bi_summary.pdf")
)
```













## Experiment 3: Bilingual POA data

Model info here. 


(ref:plot-poa-bilinguals) VOT and burst metrics of voiceless stops from bilingual speakers as a function of language (English, Spanish), place of articulation (Coronal, Bilabial). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals.

```{r, 'plot-poa-bilinguals', echo=FALSE, fig.cap="(ref:plot-poa-bilinguals)"}
knitr::include_graphics(
  here("figs", "poa_all_metrics.pdf")
)
```

**Voice-onset time.**

`r print_model_param(all_models$mod_poa_comp_vot, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_vot, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_vot, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_vot, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_vot, "language_sum:poa_sum")`

**Relative intensity.**

`r print_model_param(all_models$mod_poa_comp_ri, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_ri, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_ri, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_ri, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_ri, "language_sum:poa_sum")`

**Center of gravity.**

`r print_model_param(all_models$mod_poa_comp_cog, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_cog, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_cog, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_cog, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_cog, "language_sum:poa_sum")`

**Kurtosis.**

`r print_model_param(all_models$mod_poa_comp_kt, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_kt, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_kt, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_kt, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_kt, "language_sum:poa_sum")`

**Standard deviation.**

`r print_model_param(all_models$mod_poa_comp_sd, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_sd, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_sd, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_sd, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_sd, "language_sum:poa_sum")`

**Skewness.**

`r print_model_param(all_models$mod_poa_comp_sk, "language_sum")`  

`r print_model_param(all_models$mod_poa_comp_sk, "poa_sum")`  

`r print_model_param(all_models$mod_poa_comp_sk, "f1_std")`  

`r print_model_param(all_models$mod_poa_comp_sk, "f2_std")`  

`r print_model_param(all_models$mod_poa_comp_sk, "language_sum:poa_sum")`




(ref:plot-poa-bilinguals-summary) Posterior medians &plusmn; 95% and 66% credible intervals for VOT and burst metrics of bilingual voiceless bilabial and coronal stops.

```{r, 'plot-poa-bilinguals-summary', echo=FALSE, fig.cap="plot-poa-bilinguals-summary"}
knitr::include_graphics(
  here("figs", "poa_summary.pdf")
)
```








# Discussion

## Summary of findings

## Interpretation and implications


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
