---
author: 
  - name          : "Miquel Simonet"
    affiliation   : "1"
    corresponding : yes
    address       : "Postal address"
    email         : "my@email.com"
  - name          : "Joseph V. Casillas"
    affilitation  : "2"
  - name          : "Alex Aldrich"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Arizona"
  - id            : "2"
    institution   : "Rutgers University"
  - id            : "3"
    institution   : "Concordia College"

keywords          : ""
wordcount         : ""

bibliography      : ["r-references.bib"]

appendix:
  - "appendix_a.Rmd"
  - "appendix_b.Rmd"
  - "appendix_c.Rmd"
  - "appendix_d.Rmd"

floatsintext      : yes
figurelist        : yes
tablelist         : yes
footnotelist      : yes
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r 'doc-libs', include = FALSE}
library("papaja")
library("officer")
library("flextable")
library("brms")
```

```{r 'setup', include=FALSE}
set.seed(42)
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      cache.extra = knitr::rand_seed)
```

```{r, 'load-all'}
source(here::here("scripts", "r", "production", "99_small_data.R"))
```

# Overview

## Bayesian data analysis

This study employs Bayesian Data Analysis for quantitative inferential 
statistics. 
Specifically, this implies that we use Bayesian *credible intervals* to draw 
statistical inferences.
A Bayesian model calculates a posterior distribution, i.e., a distribution of 
plausible parameter values, given the data, a data-generating model, and any 
prior information we have about those parameter values. 
Posterior distributions are computationally costly. 
For this reason, we use the Hamiltonian Markov Chain Monte Carlo algorithm 
to obtain a sample that incldues thousands of values from the posterior 
distribution. 
In practical terms, what this means is that we do not calculate a single 
point estimate for an effect $\beta$, but rather we draw a sample of 4,000 
plausible values for $\beta$. 
This allows us to quantify our uncertainty regarding $\beta$ by summaryzing 
the distribution of those values.
We will use 3 statistics to describe the posterior distribution: (1) the mean, 
(2) a credible interval, and (3) the posterior probability that the effect is
greater than 0.
The mean provides a point estimate for the distribution, the 95% credible 
interval provides bounds for the effect, and the posterior probability that 
the effect is greater than zero further quantifies our uncertainty.
For example, if a hypothesis states that $\beta > 0$, we judge there to be 
*compelling evidence* for this hypothesis if the mean point estimate is 
a positive number, is (by a reasonably clear margin) not included in the 95% 
credible interval of $\beta$, and the posterior $P(\beta > 0)$ is close to 
one.

These three statistics provide an intuitive interpretation of any given 
effect. 
For instance, consider a case in which the posterior mean of $\beta$ is 100 
and the 95% credible interval is [40, 160].
The interval tells us that we can be 95% certain the *true* value of $\beta$ 
is between 40 and 160, given the data, our model, and our prior information. 
Furthermore, the interval allows us to specify areas of uncertainty. 
In this example, we can conclude that the effect is almost certain to be 
positive. 
The lower interval value of 40 tells us that 95% of the plausible values are 
greater than 40. 
We also note that the interval covers a wide range of values, thus we also 
conclude that we are not very certain about the size of the effect. 
This type of interepretation is not possible under a frequentist paradigm.

## About this document

This document was written in RMarkdown using `papaja` [@R-papaja] and serves 
as a project report for our research group.
From this point foward, the document is written as if it were the results 
section of a future manuscript.
This implies that it is written in a way that allows it to be copy and pasted 
into the actual manuscript once it is available.


\pagebreak




# Method

## Participants

The data include `r exp_n` participants from 3 populations: monolingual 
English speakers, monolingual Spanish speakers, Bilingual English-Spanish 
speakers.

### Monolingual English speakers

The study ingludes `r mono_en_n` monolingual English speakers from the U.S. 
Southwest. 
Age and other info here. 

### Monolingual Spanish speakers

The monolingual Spanish speakers come from Majorca, Spain. 
There are `r mono_sp_n` of them. 
Age and other info here. 

### Bilingual speakers

The English-Spanish bilinguals (n = `r bi_coronals_n + bi_poa_n`) are from 
Southern Arizona, Northern Mexico. 
Age, BLP and other info. 
Note there are two samples from this population. 
The coronal dataset includes `r bi_coronals_n` speakers and the bilabial 
dataset includes `r bi_poa_n` speakers. 

## Metrics

F1/F2, voice onset time, relative intensity, center of gravity, standard 
deviation, skewness, kurtosis

## Procedure

Decide if we will present 3 separate experiments with 3 different methods 
sections.

## Statistical analyses

All analyses were conducted in `R` [@R-base, `r r_version`]. 
We use Bayesian multilevel models fitted in `Stan` using `brms` [@R-brms_a; @R-brms_b, `r brms_version`].
Bayesian Data Analysis (BDA) has become a welcome alternative to frequentist 
statistical analysis. 
See XXX and XXX for tutorials and in depth explanations related BDA in the 
psychological and speech sciences. 
For all models, the criterion was standardized, or converted to z-scores, 
in order to facilitate comparibility between metrics. 
Continuous predictors were also standardized and categorical predictors were 
sum-to-zero coded. 
Thus for all models the intercept represents the outcome variable at the 
average state. 
We use regularizing, weakly informative priors in all models (specifics below) 
with 4,000 iterations (2,000 warmup) running on 16 processing cores. 
We quantify our uncertainty regarding a given effect by reporting point 
estimates derived from the posterior predictive distribution, including the 
95% Highest Density Credible Intervals (HDI). 
Additionally, we assume a negligible effect size of &plusmn; 0.1 (CITE COHEN) 
in order to establish a Region of Practical Equivalence (ROPE), for which we 
assess the proportion of the HDI that falls within this interval.
Finally, we report the Maximum Probability of Effect (MPE), or the Probability 
of Direction, as the proportion of the posterior distribution that is of the 
median's sign. 
We assume there to be compelling evidence for a given effect when the HDI of 
the posterior distribution does not contain 0 nor fall within the ROPE by a 
reasonable margin and the MPE is greater than 97%. 




# Results

The results are divided into 4 subsections dealing with (1) phonetic 
development over time, (2) learning trajectories, (3) comparisons with native 
bilinguals, and (4) individual differences. 




## Experiment 0: Vowels

```{r, 'plot-monolingual-vowels', echo=FALSE, fig.cap="F1 and F2 of /a/ from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 95% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "vowel_all_metrics.pdf")
)
```

```{r, 'plot-monolingual-vowels-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "vowel_summary.pdf")
)
```




## Experiment 1: Monolinguals


```{r, 'plot-monolinguals', echo=FALSE, fig.cap="VOT and burst metrics of coronal stops (/d/, /t/) from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "mono_all_metrics.pdf")
)
```

```{r, 'plot-monolinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "mono_summary.pdf")
)
```






## Experiment 2: Bilinguals

```{r, 'plot-bilinguals', echo=FALSE, fig.cap="VOT and burst metrics of coronal stops (/d/, /t/) from bilingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "bi_all_metrics.pdf")
)
```

```{r, 'plot-bilinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "bi_summary.pdf")
)
```






## Experiment 3: Bilingual POA data

```{r, 'plot-poa-bilinguals', echo=FALSE, fig.cap="VOT and burst metrics of voiceless stops from bilingual speakers as a function of language (English, Spanish), place of articulation (Coronal, Bilabial). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "poa_all_metrics.pdf")
)
```

```{r, 'plot-poa-bilinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "poa_summary.pdf")
)
```







# Results

# Discussion


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
