---
author: 
  - name          : "Miquel Simonet"
    affiliation   : "1"
    corresponding : yes
    address       : "1423 E. University Blvd. Rm. 594 Modern Languages Building, Tucson, Arizona 85721"
    email         : "simonet@email.arizona.edu"
  - name          : "Joseph V. Casillas"
    affilitation  : "2"
  - name          : "Alex Aldrich"
    affiliation   : "3"

affiliation:
  - id            : "1"
    institution   : "University of Arizona"
  - id            : "2"
    institution   : "Rutgers University"
  - id            : "3"
    institution   : "Concordia College"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  Here. 


keywords          : "Spanish, Coronal stops, Spectral moments"
wordcount         : ""

bibliography      : 
  - "r-references.bib"
  - "stats-references.bib"

appendix:
  - "appendix_a.Rmd"
  - "appendix_b.Rmd"
  - "appendix_c.Rmd"
  - "appendix_d.Rmd"
  - "appendix_e.Rmd"

floatsintext      : yes
figurelist        : yes
tablelist         : yes
footnotelist      : yes
linenumbers       : yes
mask              : no
draft             : yes

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_word
---

```{r 'doc-libs', include = FALSE}
library("papaja")
library("officer")
library("flextable")
library("brms")
```

```{r 'setup', include=FALSE}
set.seed(42)
knitr::opts_chunk$set(echo = F, message = F, warning = F, 
                      cache.extra = knitr::rand_seed)
```

```{r, 'load-all'}
source(here::here("scripts", "r", "production", "99_small_data.R"))
```



# Method

## Participants

The data include `r exp_n` participants from 3 populations: monolingual 
English speakers, monolingual Spanish speakers, and bilingual Spanish-English 
speakers.
All participants were females between the ages of 18 and 23. 
The monolingual English speakers were recorded in English and the monolingual 
Spanish speakers were recorded in Spanish. 
The Spanish-English bilinguals were recorded in both of their languages.

### Monolingual English speakers

The study includes `r mono_en_n` monolingual English speakers.
They were undergraduate students at the University of Arizona, born and raised 
in the US Southwest. 
The English speakers were functionally monolingual, though they reported having 
taken introductory Spanish courses. 
They were not able to maintain a basic conversation in Spanish.
All of the participants in this group reported English as their native 
language and verified not having been exposed to any other languages while 
growing up.


### Monolingual Spanish speakers

The monolingual Spanish group comprised `r mono_sp_n` speakers that were 
recruited from the *Universitat de les  Illes Balears* campus community and 
were born and raised on the island of Majorca, Spain.
They reported that, although they had studied some English in Spain, they were 
not able to maintain a basic conversation in this language. 
The participants of this group also speak Catalan. 
Importantly, there are no reported differences in the phonetic realization of 
voice timing between the Spanish and Catalan, nor are there place difference 
between the coronal stops. 

### Bilingual speakers

The English-Spanish bilinguals (n = `r bi_coronals_n + bi_poa_n`) came from 
Southern Arizona and Northern Mexico. 
There are two samples from this population. 
The coronal dataset includes `r bi_coronals_n` speakers and the bilabial 
dataset includes `r bi_poa_n` speakers. 
The Spanish-English bilinguals were undergraduate students at the University 
of Arizona in Tucson, Arizona. 
The bilinguals were brought up by Spanish-speaking families and were schooled  mostly in English. 
They reported using English and Spanish daily, both in the classroom as well 
as with their friends and relatives.


## Metrics

F1/F2, voice onset time, relative intensity, center of gravity, standard 
deviation, skewness, kurtosis

## Procedure

Decide if we will present 3 separate experiments with 3 different methods 
sections.

## Statistical analyses

All analyses were conducted in `R` [@R-base, `r r_version`]. 
We use Bayesian multilevel models fitted in `Stan` using `brms` [@R-brms_a; @R-brms_b, `r brms_version`].
Bayesian Data Analysis (BDA) has become a welcome alternative to frequentist 
statistical analysis. 
See @van2014bayesian and @vasishth2018bayesian for tutorials and in depth 
explanations related BDA in the psychological and speech sciences.
For all models, the criterion was standardized, or converted to z-scores, 
in order to facilitate comparibility between metrics. 
Continuous predictors were also standardized and categorical predictors were 
sum-to-zero coded. 
Thus for all models the intercept represents the outcome variable at the 
grand mean in standardized units. 
We used regularizing, weakly informative priors in all models (specifics below) 
with 4,000 iterations (2,000 warmup) running on 16 processing cores. 
We quantify our uncertainty regarding a given effect by reporting point 
estimates derived from the posterior predictive distribution, including the 
95% Highest Density Credible Intervals (HDI). 
Additionally, we assume a negligible effect size of &plusmn; 0.1 [@cohen1988statistical; @cohen2013statistical; @kruschke2018rejecting] 
in order to establish a Region of Practical Equivalence (ROPE), for which we 
assess the proportion of the HDI that falls within this interval.
Finally, we report the Maximum Probability of Effect (MPE), or the Probability 
of Direction, as the proportion of the posterior distribution that is of the 
median's sign. 
We assume there to be compelling evidence for a given effect when the HDI of 
the posterior distribution does not contain 0 nor fall within the ROPE by a 
reasonable margin and the MPE is close to 1. 




# Results

The results are divided into 4 subsections dealing with (1) F1/F2, (2) 
monolingual data, (3) bilingual data, and (4) POA data. 




## Experiment 0: Vowels

The first analysis scrutinezed F1 and F2 of the vowel following the coronal 
stops. 
The purpose of this analysis was to determine if the low /a/ and /ae/ vowels 
of Spanish and English, respectively, differed from each other. 
This analysis was taken as a precautionary measure with the objective of 
determining whether or not coarticulation effects may be present in the stop 
metrics due to the possible different spectral envelopes of the next 
segment. 

The F1 and F2 data were analayzed using separate Bayesian multilevel models.
The criterion (F1 or F2) were standardized and modeled as a function of 
language (English, Spanish), segment (/d/, /t/), and item repetition. 
Language and segment were sum coded (-1, 1). 
The random effects structure included by-subject intercepts with 
random slopes for segment and item repetition, as well as by-item intercepts 
with a random slope for item repetition. 
The model included weakly informative priors with the mean centered at 0 and 
a standard deviation of 2. 

The model found weak evidence for a language effect on F1 
`r print_model_param(all_models$mod_f1_mono_full, "language_sum")`
and F2 
`r print_model_param(all_models$mod_f2_mono_full, "language_sum")`. 
Fifteen and eighteen percent of the HDI's fell within the specified ROPE for 
F1 and F2, respectively. 
The probability that &beta; was greater than 0 was .979 for F1 and .958 for F2. 
Together, the point estimates suggest the spectral centroid of the Spanish 
vowel was slightly higher and more posterior with respect to the English vowel.
There was no evidence for a segment effect on F1 
`r print_model_param(all_models$mod_f1_mono_full, "phon_sum")`, 
nor on F2 
`r print_model_param(all_models$mod_f2_mono_full, "phon_sum")`. 
In both cases 100% of the HDI fell within the ROPE and the probability that 
&beta; was less than 0 was .717 for F1 and .555 for F2. 
Given the possibility of a small effect of language on the spectal envelope, 
standardized F1 and F2 were used in subsequent analyses of the coronal stops. 
Figure 1 plots the F1 &#215; F2 data and Figure 2 plots the model summary. 
A complete table of the F1 and F2 models is available in Appendix A. 


```{r, 'plot-monolingual-vowels', echo=FALSE, fig.cap="F1 &#215; F2 of /a/ from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 95% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "vowel_all_metrics.pdf")
)
```

```{r, 'plot-monolingual-vowels-summary', echo=FALSE, fig.cap="Posterior medians  &plusmn; 95% and 66% credible intervals for F1 and F2 data."}
knitr::include_graphics(
  here("figs", "vowel_summary.pdf")
)
```




## Experiment 1: Monolinguals


```{r, 'plot-monolinguals', echo=FALSE, fig.cap="VOT and burst metrics of coronal stops (/d/, /t/) from monolingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "mono_all_metrics.pdf")
)
```

```{r, 'plot-monolinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "mono_summary.pdf")
)
```






## Experiment 2: Bilinguals

```{r, 'plot-bilinguals', echo=FALSE, fig.cap="VOT and burst metrics of coronal stops (/d/, /t/) from bilingual speakers as a function of language (English, Spanish). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "bi_all_metrics.pdf")
)
```

```{r, 'plot-bilinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "bi_summary.pdf")
)
```






## Experiment 3: Bilingual POA data

```{r, 'plot-poa-bilinguals', echo=FALSE, fig.cap="VOT and burst metrics of voiceless stops from bilingual speakers as a function of language (English, Spanish), place of articulation (Coronal, Bilabial). Transparent points represent raw data. Solid points indicate posterior means &plusmn; 99% and 80% credible intervals."}
knitr::include_graphics(
  here("figs", "poa_all_metrics.pdf")
)
```

```{r, 'plot-poa-bilinguals-summary', echo=FALSE, fig.cap="Model plot."}
knitr::include_graphics(
  here("figs", "poa_summary.pdf")
)
```







# Results

# Discussion

effectSize <- (mu1 - mu2) / sqrt((sigma1^2 + sigma2^2) / 2)


\newpage

# References
```{r create_r-references}
r_refs(file = "r-references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id = "refs"></div>
\endgroup
